Progress Report on One Simulated Dataset

### 2023-11-27 Docs/ProgressReportSim.Rmd
simulation result summary

library(here)
library(tidyverse)
library(ggpubr)
library(refund)
library(lme4)
library(ROCR)
library(GLMMadaptive)
library(kableExtra)
library(knitr)
library(mvtnorm)
library(mgcv)
library(splines)
theme_set(theme_minimal())

# load simulation results
load(here("Data/SimOutput_fGFPCA.RData"))
```

@@

@@

```{r pred_debias_unscale, results='hide'}
# approximation process
# container
pred_list <- list()
converge_vec <- matrix(NA, N_test, 4)
# score <- array(NA, dim=c(length(test_id), K, 4))
# dimensions: subject, eigenfunction, maximum observation time

@@

    # into a list
    MyData <- list(K=K, 
                   X=efunctions_new[1:max_rid, ], 
                   mon.names=mon.names,
                   parm.names=parm.names, 
                   pos.xi=pos.xi, 
                   y=df_it$Y, 
                   tao=diag(new_lambda), f0=new_mu[1:max_rid])

@@

p1<-df_pred %>%
  filter(id %in% rand_test_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "Unscaled, debiased eigenvalues")

@@

```{r pred_debias_scale, results='hide'}
## With debiased eigenvalues, also scale eigenvaleus and eigenfuncitons
# container
t1 <- Sys.time()
pred_list3 <- list()
converge_vec2 <- matrix(NA, N_test, 4)
# per subject
for(i in seq_along(test_id)){
  df_i <- test_df %>% filter(id==test_id[i])

    # into a list
    MyData <- list(K=K,
                   X=efunctions_scaled[1: max_rid, ], 
                   mon.names=mon.names,
                   parm.names=parm.names, 
                   pos.xi=pos.xi, 
                   y=df_it$Y, 
                   tao=diag(scaled_lambda), f0=new_mu[1:max_rid])

    # fit laplace approximation
    Fit <- LaplaceApproximation(Model, parm = rep(0, K), Data=MyData, Method = "NM", Iterations = 1000,
                                CovEst = "Identity")
    converge_vec2[i, which(c(0.2, 0.4, 0.6, 0.8)==tmax)] <- Fit$Converged
    score <- Fit$Summary1[, "Mode"]

    # prediction
    eta_pred_out <- new_mu+efunctions_scaled%*%score
    df_i[ , paste0("pred", tmax)] <- eta_pred_out[,1]
  }

  pred_list3[[i]] <- df_i
}
t2 <- Sys.time()
# t2-t1
# mean(converge_vec2)
```

```{r}
df_pred3 <- bind_rows(pred_list3)

df_pred3$pred0.2[df_pred3$t<=0.2] <- NA
df_pred3$pred0.4[df_pred3$t<=0.4] <- NA
df_pred3$pred0.6[df_pred3$t<=0.6] <- NA
df_pred3$pred0.8[df_pred3$t<=0.8] <- NA
# overview of predicted track
p3 <- df_pred3 %>%
  filter(id %in% rand_test_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "Scaled, unbiased eigenvalues")

```{r}
t_1sim <- (t_local_glmm+t_fpca+t_debias+t_pred)
```

I would like to see if the following operations can improve predictive performance:

1. Debiasing eigenvalues from FPCA
2. Scale the eigenfunctions and debiased eigenvalues (from mgcv::bam) so that they are close to the true values for data generation

```{r, fig.height=5, fig.width=10}
ggarrange(p1, p3, nrow = 1, common.legend = T)
```

- Time spent on one iteration: `r round(t_1sim, 2)` seconds
- Debiasing eigenvalues helped with prediction. Scaling eigenfunctiosn and eigenvalues, on the other hand, ended up being more complicated.

## ISE

@@

```{r}
# calcualte ISE
err3 <- df_pred3 %>%
  mutate(err1 = (pred0.2-eta_i)^2,
         err2 = (pred0.4-eta_i)^2,
         err3 = (pred0.6-eta_i)^2,
         err4 = (pred0.8-eta_i)^2) %>%
  select(id, t, starts_with("err"))

ise <- full_join(tb1, tb3, by = "Window")
colnames(ise) <-c("Window",  rep(seq(0.2, 0.8, by =0.2), 2))
ise %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "= 1, "Unscaled" = 4, "Scaled" = 4)) %>%
  add_header_above(c(" "=1, "Maximum observation time" = 8))

## AUC

@@

```{r}
# calcualte AUC
auc_df3 <- df_pred3 %>%
  mutate(window = cut(t, breaks = seq(0, 1, by = 0.2), include.lowest = T)) %>%
  select(Y, starts_with("pred"), window)

auc_mat3 <- matrix(NA, length(pred_window), length(obs_track))

for(j in seq_along(obs_track)){
  for(i in (j+1):5){
    this_Y <- auc_df3$Y[auc_df3$window==pred_window[i]]
    this_pred <- auc_df3[auc_df3$window==pred_window[i],
                         paste0("pred", obs_track[j])]
    this_perf <- performance(prediction(this_pred, this_Y), measure = "auc")
    auc_mat3[i ,j] <- this_perf@y.values[[1]]
  }
}

auc <- cbind(auc_mat, auc_mat3)
rownames(auc) <- unique(auc_df1$window)
colnames(auc) <-c(rep(seq(0.2, 0.8, by = 0.2), 2))
auc %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "= 1, "Unscaled" = 4, "Scaled" = 4)) %>%
  add_header_above(c(" "=1, "Maximum observation time" = 8))
```

@@

Below I compared the predictions using scaled and unscaled parameters:

```{r, fig.height=8, fig.width=8}
par(mfrow = c(2, 2))
plot(df_pred$pred0.2, df_pred3$pred0.2, pch=20, cex = 0.5,
     main = "Up to 0.2", xlab = "Unscaled", ylab = "Scaled")
plot(df_pred$pred0.4, df_pred3$pred0.4, pch=20, cex = 0.5,
     main = "Up to 0.4", xlab = "Unscaled", ylab = "Scaled")
plot(df_pred$pred0.6, df_pred3$pred0.6, pch=20, cex = 0.5,
     main = "Up to 0.6", xlab = "Unscaled", ylab = "Scaled")
plot(df_pred$pred0.8, df_pred3$pred0.8, pch=20, cex = 0.5,
     main = "Up to 0.8", xlab = "Unscaled", ylab = "Scaled")
```

@@

## GLMMadaptive

- Here we can fit a model with random intercept and slope for time

$$g(E(Y_i(t))) = \beta_0+\beta_1t+b_{i0}+b_{i1}t$$
- But this model is obviously too simple for the data generation scheme

@@

@@

- Fit on all 500 subjects take too long. I reduced the sample size to 10 
- In fact, 10 samples took more than 30 minutes to fit the model. Therefore, I further reduce the number of observations for each subject. I'll start with only the 1/10 of the time track (100 observations), take observations at bin midpoints 5, 15, 25, ...., 995. This is to preserve the non-linearity of latent function when reducing grid density

$$g(E(Y_i(t))) = \sum_{k=1}^5\zeta_{k}B_k(t)+\sum_{l=1}^4\xi_{il}\phi_l(t)$$

```{r}
# subset 10 subjects from training dataset 
df_train_sub <- train_df %>% filter(id %in% 1:10 & sind==bin)
df_train_sub$id <- droplevels(df_train_sub$id)
# levels(df_train_sub$id)
# intersect(t, unique(df_train_sub$t))
# View(df_train_sub)
```

- Below I construct basis functions for the fixed effect/random intercept. Let's use five basis functions to start with. I believe this is far less flexible than the FPCA model, but it still takes a really long time to run. It is actually not realistic to repeat 500 times.

PS 50 subjects took 1.06 hours to fit with the same model structure.

```{r}
# construct design matrix for the fixed effect intercept
t_sub <- unique(df_train_sub$t)
# fix_xmat <- smoothCon(s(t, bs="cr", k=5), data = data.frame(t=t_sub))
fix_xmat <- ns(t_sub, df=5)
# dim(fix_xmat[[1]]$X)

# basis functions
fix_xmat <- data.frame(t=t_sub, 
                       fix_xmat)
colnames(fix_xmat) <- c("t", paste0("B", 1:5))

df_train_sub <- df_train_sub %>% left_join(fix_xmat, by = "t")
```

- Below I construct basis functions for the individual random effect (also a functinal intercept), with four basis functions to start with.

```{r}
# construct design matrix for the random effect
rand_xmat <- smoothCon(s(t, bs="cr", k=4), data = data.frame(t=t_sub))
# rand_xmat <- ns(t_sub, df = 4)

# dim(rand_xmat[[1]]$X) # 5 basis functions for each of the 50 subjects
rand_xmat <- data.frame(t = t, rand_xmat[[1]]$X)

df_train_sub <- df_train_sub %>%
  left_join(rand_xmat, by = c("t"))
```

- Now I'll fit the model and record the time spent on model fitting

The model fitting took about 12 minutes at this data scale, which obviously is not realistic for 500 simulated datasets.

Also tried increasing number of basis functions, which ended up with "computationally singular". In fact, this method does not seem very numerically stable.

```{r, eval=TRUE, class.source = "fold-show"}
# model fit
t1 <- Sys.time()
fit_adglmm <- mixed_model(fixed = Y ~ 0+B1+B2+B3+B4+B5, 
                          random = ~X1+X2+X3+X4 | id,
                          data =  df_train_sub, family = binomial())
t2 <- Sys.time()
# t2-t1

@@

- But anyways, let's try to use this method for prediction. Let's do that for 10 new test subjects.

```{r}
df_test_sub <- test_df %>% filter(id %in% 1:10 & sind %in% mid)
df_test_sub$id <- droplevels(df_test_sub$id)
# levels(df_test_sub$id)

# basis functions for both fixed and random effects should be shared across sample
df_test_sub <- df_test_sub %>% left_join(fix_xmat, by = "t") %>%
  left_join(rand_xmat, by = "t")

# given observation from 0.2, 0.4, 0.6, 0.8, predict the rest
# this step in fact is very fast
pred_adglmm1 <- predict(fit_adglmm, 
                      newdata = df_test_sub %>% filter(t<=0.2),
                      newdata2 =  df_test_sub %>% filter(t>0.2),
                      type = "subject_specific", type_pred = "link",
                      return_newdata = TRUE)

pred_adglmm2 <- predict(fit_adglmm, 
                      newdata = df_test_sub %>% filter(t<=0.4),
                      newdata2 =  df_test_sub %>% filter(t>0.4),
                      type = "subject_specific", type_pred = "link",
                      return_newdata = TRUE)

pred_adglmm3 <- predict(fit_adglmm, 
                      newdata = df_test_sub %>% filter(t<=0.6),
                      newdata2 =  df_test_sub %>% filter(t>0.6),
                      type = "subject_specific", type_pred = "link",
                      return_newdata = TRUE)

pred_adglmm4 <- predict(fit_adglmm, 
                      newdata = df_test_sub %>% filter(t<=0.8),
                      newdata2 =  df_test_sub %>% filter(t>0.8),
                      type = "subject_specific", type_pred = "link",
                      return_newdata = TRUE)
```

```{r}
pred_adglmm_all <- df_test_sub %>% 
  left_join(pred_adglmm1$newdata2 %>% select(id, t, pred), by = c("id", "t")) %>% 
  left_join(pred_adglmm2$newdata2 %>% select(id, t, pred), by = c("id", "t")) %>%
  left_join(pred_adglmm3$newdata2 %>% select(id, t, pred), by = c("id", "t")) %>%
  left_join(pred_adglmm4$newdata2 %>% select(id, t, pred), by = c("id", "t"))

colnames(pred_adglmm_all) <- c(colnames(pred_adglmm_all)[1:14], "pred1", "pred2", "pred3", "pred4")
```

```{r}

pred_adglmm_all %>%
  filter(id %in% 1:4) %>%
  mutate_at(vars(eta_i, pred1, pred2, pred3, pred4), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred1, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred2, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred3, col = "0.5"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred4, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "GLMMadaptive + spline basis prediction")

@@

# Notes

Perhaps I should try to verify my thoughts about laplace approximation by different generation scheme. For example, what if the true eigenvalues are on larger scale?

- Well, after increasing the scale of the true eigenvalues, the convergence rate did get closer between scaled and unscaled parameter. Also reversed as I expected.
- But on second though, A more effecient way is to calcualted ISE and AUC stratified by convergence status

>>>>

library(here)
library(tidyverse)
library(ggpubr)
library(refund)
library(lme4)
library(ROCR)
library(GLMMadaptive)
library(kableExtra)
library(knitr)
library(mvtnorm)
library(mgcv)
library(splines)
library(LaplacesDemon)
theme_set(theme_minimal())

# load simulation results
# load(here("Data/SimOutput_fGFPCA.RData"))
```

@@

```{r}
# functions
source(here("Code/GLMM-FPCA.R")) 
source(here("Code/OutSampBayes.R"))
```

@@

```{r pred_debias_scale, results='hide'}
# approximation process
# container
pred_list <- list()
converge_vec <- matrix(NA, N_test, 4)
# score <- array(NA, dim=c(length(test_id), K, 4))
# dimensions: subject, eigenfunction, maximum observation time

@@

    # into a list
    MyData <- list(K=K, 
                   X=efunctions_scaled[1:max_rid, ], 
                   mon.names=mon.names,
                   parm.names=parm.names, 
                   pos.xi=pos.xi, 
                   y=df_it$Y, 
                   tao=diag(scaled_lambda), f0=new_mu[1:max_rid])

@@

p1<-df_pred %>%
  filter(id %in% rand_test_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "fGFPCA")

@@

@@

@@

@@

@@

## GLMMadaptive linear time effect

- Here we can fit a model with random intercept and slope for time. It is doable on 500 datasets, but obviously too simple for the data generation scheme

$$g(E(Y_i(t))) = \beta_0+\beta_1t+b_{i0}+b_{i1}t$$

@@

```{r}
# out-of-sample prediction
pred_lin_glmm <- list(
  predict(fit_adglmm, 
          newdata = test_df %>% filter(t<=0.2),
          newdata2 =  test_df %>% filter(t>0.2),
          type = "subject_specific", type_pred = "link",
          return_newdata = TRUE)$newdata2,

  predict(fit_adglmm, 
          newdata = test_df %>% filter(t<=0.4),
          newdata2 =  test_df %>% filter(t>0.4),
          type = "subject_specific", type_pred = "link",
          return_newdata = TRUE)$newdata2,

  predict(fit_adglmm, 
          newdata = test_df %>% filter(t<=0.6),
          newdata2 =  test_df %>% filter(t>0.6),
          type = "subject_specific", type_pred = "link",
          return_newdata = TRUE)$newdata2,

  predict(fit_adglmm, 
          newdata = test_df %>% filter(t<=0.8),
          newdata2 =  test_df %>% filter(t>0.8),
          type = "subject_specific", type_pred = "link",
          return_newdata = TRUE)$newdata2
  )
```

```{r}
pred_lin_glmm <- lapply(pred_lin_glmm, 
                        function(df){df %>% select(id, t, pred)})

pred_lin_glmm_df <- test_df %>%
  left_join(pred_lin_glmm[[1]], by = c("id", "t")) %>%
  rename(pred0.2 = pred) %>%
  left_join(pred_lin_glmm[[2]], by = c("id", "t")) %>%
  rename(pred0.4 = pred) %>%
  left_join(pred_lin_glmm[[3]], by = c("id", "t")) %>%
  rename(pred0.6 = pred) %>%
  left_join(pred_lin_glmm[[4]], by = c("id", "t")) %>%
  rename(pred0.8 = pred) 
```

```{r}
p2 <- pred_lin_glmm_df %>%
  filter(id %in% rand_test_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "GLMMadaptive")
```

```{r, fig.height=5, fig.width=10}
ggarrange(p1, p2, nrow = 1, common.legend = T)
```

```{r}
# calcualte ISE
err2 <- pred_lin_glmm_df %>%
  mutate(err1 = (pred0.2-eta_i)^2,
         err2 = (pred0.4-eta_i)^2,
         err3 = (pred0.6-eta_i)^2,
         err4 = (pred0.8-eta_i)^2) %>%
  select(id, t, starts_with("err"))

err2$window = cut(err2$t, breaks = seq(0, 1, by = 0.2), include.lowest = T)
# table(err1$window)
err2 <- split(err2, f = err2$window)

tb2 <- lapply(err2, function(x){
  x %>%  group_by(id) %>% 
    summarise_at(vars(err1, err2, err3, err4), sum) %>%
    summarise_at(vars(err1, err2, err3, err4), mean)
}) %>% bind_rows(.id = "Window")
```

ise <- full_join(tb1, tb2, by = "Window")
colnames(ise) <-c("Window",  rep(seq(0.2, 0.8, by =0.2), 2))
ise %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"",
        caption = "ISE") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "= 1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "=1, "Maximum observation time" = 8))

```{r}
# calcualte AUC
auc_df2 <- pred_lin_glmm_df %>%
  mutate(window = cut(t, breaks = seq(0, 1, by = 0.2), include.lowest = T)) %>%
  select(Y, starts_with("pred"), window)

auc_mat2 <- matrix(NA, length(pred_window), length(obs_track))

for(j in seq_along(obs_track)){
  for(i in (j+1):5){
    this_Y <- auc_df2$Y[auc_df2$window==pred_window[i]]
    this_pred <- auc_df2[auc_df2$window==pred_window[i],
                         paste0("pred", obs_track[j])]
    this_perf <- performance(prediction(this_pred, this_Y), measure = "auc")
    auc_mat2[i ,j] <- this_perf@y.values[[1]]
  }
}

auc <- cbind(auc_mat, auc_mat2)
rownames(auc) <- unique(auc_df1$window)
colnames(auc) <-c(rep(seq(0.2, 0.8, by = 0.2), 2))
auc %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"",
        caption = "AUC") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "= 1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "=1, "Maximum observation time" = 8))
```

@@

Due to computation time, I took only 10 subjects for training set. In this case, the model took 30 minutes to fit. Perhaps we could do a few more iterations?

I use five basis functions for both fixed and random effects.

$$g(E(Y_i(t))) = \sum_{k=1}^5\zeta_{k}B_k(t)+\sum_{l=1}^5\xi_{il}\phi_l(t)$$

```{r}
# subset 10 subjects from training dataset 
df_train_sub <- train_df %>% filter(id %in% 1:10)
df_train_sub$id <- droplevels(df_train_sub$id)

df_test_sub <- test_df %>% filter(id %in% 1:10)
df_test_sub$id <- droplevels(df_test_sub$id)
# levels(df_test_sub$id)
```

```{r flex_adglmm, cache=T}
t1 <- Sys.time()
fit_adglmm_dense <- mixed_model(fixed = Y ~ 0+ns(t, df = 5), 
                          random = ~ 0+ns(t, df = 5) | id, 
                          data =  df_train_sub, family = binomial())
t2 <- Sys.time()
# t2-t1
```

As we have talked about from last meeting, I wanna use ns() wrapped in the GLMMadaptive::mixed_model() to see if this could speed things up. The model fit took 
`r round(t2-t1, 2)` mins to fit.

```{r}
# given observation from 0.2, 0.4, 0.6, 0.8, predict the rest
pred_nonlin_adglmm <- list(
  predict(fit_adglmm_dense, 
          newdata = df_test_sub %>% filter(t<=0.2),
          newdata2 =  df_test_sub %>% filter(t>0.2),
          type = "subject_specific", type_pred = "link",
          return_newdata = TRUE)$newdata2,
  predict(fit_adglmm_dense, 
          newdata = df_test_sub %>% filter(t<=0.4),
          newdata2 =  df_test_sub %>% filter(t>0.4),
          type = "subject_specific", type_pred = "link",
          return_newdata = TRUE)$newdata2,
  predict(fit_adglmm_dense, 
          newdata = df_test_sub %>% filter(t<=0.6),
          newdata2 =  df_test_sub %>% filter(t>0.6),
          type = "subject_specific", type_pred = "link",
          return_newdata = TRUE)$newdata2,
  predict(fit_adglmm_dense, 
          newdata = df_test_sub %>% filter(t<=0.8),
          newdata2 =  df_test_sub %>% filter(t>0.8),
          type = "subject_specific", type_pred = "link",
          return_newdata = TRUE)$newdata2
  )
```

```{r}
pred_nonlin_adglmm <- lapply(pred_nonlin_adglmm, 
                        function(df){df %>% select(id, t, pred)})

pred_nonlin_adglmm_df <- df_test_sub %>%
  left_join(pred_nonlin_adglmm[[1]], by = c("id", "t")) %>%
  rename(pred0.2 = pred) %>%
  left_join(pred_nonlin_adglmm[[2]], by = c("id", "t")) %>%
  rename(pred0.4 = pred) %>%
  left_join(pred_nonlin_adglmm[[3]], by = c("id", "t")) %>%
  rename(pred0.6 = pred) %>%
  left_join(pred_nonlin_adglmm[[4]], by = c("id", "t")) %>%
  rename(pred0.8 = pred)

# pred_nonlin_adglmm_df %>% filter(t>0.4)
```

```{r}
p1_nonlin <- pred_nonlin_adglmm_df %>%
  filter(id %in% 1:4) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8),
            function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.5"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "GLMMadaptive + spline basis prediction")

@@

```{r}
# calcualte ISE
err3 <- pred_nonlin_adglmm_df %>%
  mutate(err1 = (pred0.2-eta_i)^2,
         err2 = (pred0.4-eta_i)^2,
         err3 = (pred0.6-eta_i)^2,
         err4 = (pred0.8-eta_i)^2) %>%
  select(id, t, starts_with("err"))

For the comparing fGFPCA, if we decrease the measurement density to 100 points, a lot of numeric problems popped up in the either local models or Laplace Approximation process. It looks like when observation track is short, GLMMadaptive has a better prediction.

```{r fGFPCA}
# bin data
# J_sub <- length(unique(df_train_sub$t))
# df_train_sub$sind <- rep(1:J_sub, 10)
# t_sub <- unique(df_train_sub$t)
# bin_w <- 4

# n_bin_sub <- J_sub/bin_w # number of bins
# brks <- seq(0, J_sub, by = bin_w) # cutoff points on time index
# mid <- (brks+bin_w/2)[1:n_bin_sub] # interval mid point time index
# mid_t <- t_sub[mid] 
# 
# df_train_sub$bin <- cut(df_train_sub$sind,
#                         breaks = brks, include.lowest = T, labels = mid)
# df_train_sub$bin <- as.numeric(as.character(df_train_sub$bin))
```

```{r}
# fit fGFPCA model on the training set
t1 <- Sys.time()

## Step 2: local GLMM and estimate latent function
## on training set
bin_lst_sub <- split(df_train_sub, f = df_train_sub$bin)
df_est_latent_sub <- lapply(bin_lst_sub, function(x){pred_latent(x, n_node = 0)}) 
# boundary hitting in this process
df_est_latent_sub <- bind_rows(df_est_latent_sub) 
# range(df_est_latent_sub$eta_hat)
```

```{r}
## Step 3: FPCA
uni_eta_hat_sub <- df_est_latent_sub %>% filter(bin==sind)
mat_est_unique <- matrix(uni_eta_hat_sub$eta_hat,
                         nrow=10, 
                         ncol=n_bin, byrow = F) 
fpca_mod_sub <- fpca.face(mat_est_unique, argvals = mid_t, var=T)
```

```{r}
## Step 4: project and debias
## Projection
# values used for projection
p <- 3 # order of b splines 
knots <- 35 # number of knots (same from FPCA model)
knots_values <- seq(-p, knots + p, length = knots + 1 + 2 *p)/knots
knots_values <- knots_values * (max(mid_t) - min(mid_t)) + min(mid_t)
B <- spline.des(knots = knots_values, x = mid_t, ord = p + 1,
                outer.ok = TRUE)$design  # evaluate B-splines on binned grid
Bnew <- spline.des(knots = knots_values, x = t, ord = p + 1,
                   outer.ok = TRUE)$design  # evaluate B-splines on original grid
df_phi <- matrix(NA, J, K) 
for(k in 1:K){
  lm_mod <- lm(fpca_mod_sub$efunctions[,k] ~ B-1)
  df_phi[,k] <- Bnew %*% coef(lm_mod)
}# project binned eigenfunctions onto the original grid
```

```{r}
## debias
df_phi <- data.frame(t = t, df_phi)
colnames(df_phi) <- c("t", paste0("phi", 1:4))
df_train_sub <- df_train_sub %>% 
  select(-starts_with("phi")) %>%
  left_join(df_phi, by = "t")
train_df$id <- as.factor(train_df$id)
debias_glmm <- bam(Y ~ s(t, bs="cc", k=10)+
                     s(id, by=phi1, bs="re")+
                     s(id, by=phi2, bs="re")+
                     s(id, by=phi3, bs="re")+
                     s(id, by=phi4, bs="re"), 
                   family = binomial, data=df_train_sub, 
                   method = "fREML",
                   discrete = TRUE)
new_mu <- predict(debias_glmm, type = "terms")[1:J, 1] # extract re-evaluated mean
new_lambda <- 1/debias_glmm$sp[2:5] # extract re-evaluated lambda
# rescale
new_phi <- df_phi %>% select(starts_with("phi"))*sqrt(n_bin)
new_phi <- as.matrix(new_phi)
new_lambda <- new_lambda/n_bin
t2 <- Sys.time()

@@

```{r, results='hide'}
df_pred_sub<- df_test_sub
converge_st <- matrix(NA, 10, 4)

for(i in 1:10){
    df_i <- df_test_sub %>% filter(id==i)

      # into a list
      MyData <- list(K=K, 
                     X=new_phi[1:max_rid, ], 
                     mon.names=mon.names,
                     parm.names=parm.names, 
                     pos.xi=pos.xi, 
                     y=df_it$Y, 
                     tao=diag(new_lambda), f0=new_mu[1:max_rid])

      # fit laplace approximation
      Fit <- LaplaceApproximation(Model, parm = rep(0, K), 
                                  Data=MyData, Method = "NM", Iterations = 1000,
                                  CovEst = "Identity")
      converge_st[i, which(c(0.2, 0.4, 0.6, 0.8)==tmax)] <- Fit$Converged
      score <- Fit$Summary1[, "Mode"]

      # prediction
      eta_pred_out <- new_mu+new_phi%*%score
      df_pred_sub[df_pred_sub$id==i, paste0("pred", tmax)] <- eta_pred_out[,1]
    }
}

```{r}
df_pred_sub$pred0.2[df_pred_sub$t<=0.2] <- NA
df_pred_sub$pred0.4[df_pred_sub$t<=0.4] <- NA
df_pred_sub$pred0.6[df_pred_sub$t<=0.6] <- NA
df_pred_sub$pred0.8[df_pred_sub$t<=0.8] <- NA

p2_fgfpca <- df_pred_sub %>%
  filter(id %in% 1:4) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8),
            function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.5"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "fGFPCA")

```{r}
# calcualte ISE
err4 <- df_pred_sub %>%
  mutate(err1 = (pred0.2-eta_i)^2,
         err2 = (pred0.4-eta_i)^2,
         err3 = (pred0.6-eta_i)^2,
         err4 = (pred0.8-eta_i)^2) %>%
  select(id, t, starts_with("err"))

err4$window = cut(err4$t, breaks = seq(0, 1, by = 0.2), include.lowest = T)
# table(err1$window)
err4 <- split(err4, f = err4$window)

tb4 <- lapply(err4, function(x){
  x %>%  group_by(id) %>% 
    summarise_at(vars(err1, err2, err3, err4), sum) %>%
    summarise_at(vars(err1, err2, err3, err4), mean)
}) %>% bind_rows(.id = "Window")
```

ise <- full_join(tb4, tb3, by = "Window")
colnames(ise) <-c("Window",  rep(seq(0.2, 0.8, by =0.2), 2))
ise %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"",
        caption = "ISE") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "= 1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "=1, "Maximum observation time" = 8))
```

```{r, fig.height=5, fig.width=10}
ggarrange(p1_nonlin, p2_fgfpca, nrow = 1, common.legend = T)
```

## Random thouds
